Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-06 paper_jasa.tex
Finished re-working the contaminated data section

@@

The first two simulation experiments are based on normal error model which satisfies the conditions for uniformly most powerful (UMP) test. To study how the proposed visual inference work for a scenario where the data are not really coming from a normal model we design a contaminated data experiment. To generate the contaminated data set of size $n=100$ we used the following model

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ which produced actual data size 115, $\alpha=0$ and $\lambda=10$ . $\sigma$ is chosen such that overall error standard deviation of the data is 5 which is approximately 3.5. One such plot of the contaminated data for  $\beta$=5 is shown in figure \ref{fig:cont_dat}.  We fit model (Equation \eqref{multi}) with $p=1$  and intercept $\beta_0=0$ to this contaminated data to test the hypothesis whether slope is zero or not.

{\bf Steps to obtain lineup}
\begin{enumerate}
	\item For a given parameter setting obtain actual data set
	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
\end{enumerate}

One lineup for slope $\beta=0.4$ is shown in figure \ref{fig:test_contaminated}. For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups. each subject got at least 10 lineups to evaluate. Based on the slope parameter we group lineups into six groups with difficulty = 0 for largest slope and difficulty = 5 for the smallest slope. Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.

>>>>

\green{Discuss the UMP test, and give the reference for it here.}

@@

The first two simulation experiments use data generated with a normal error model, satisfying the conditions for uniformly most powerful (UMP) test procedures. In these situations there exists a test, and there would, in general, be no need to use visual inference. The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the UMP, will compare favorably with it. This third simulation is closer to the mark for the purpose of visual inference. The assumptions for the UMP are violated, by contaminating the data. The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.  The data is generated from the following model:

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????}. Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  A linear model (Equation \eqref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

Five replications of actual data sets for each slope value shown in Table \ref{tbl:experiment_params}, producing 30 lineups. Six groups of difficulty level based on slope are defined, difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.  \green{This is the reverse of the previous experiment}  Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.

%{\bf Steps to obtain lineup}
%\begin{enumerate}
%	\item For a given parameter setting obtain actual data set
%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
%\end{enumerate}

An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? Plot 7 is the actual data plot.
